### 2018-02-21 reports/initial-proposal/initial-proposal.tex
Initial commit

%        File: initial-proposal.tex
%     Created: Tue Feb 20 10:00 PM 2018 E
% Last Change: Tue Feb 20 10:00 PM 2018 E
%
\documentclass[a4paper]{amsart}
\begin{document}

\title{APMTH 221 Final Project Proposal}
\author{Jiafeng (Kevin) Chen \and Francisco Rivera}
\date{\today}

\maketitle

\section{Problem Statement}

\end{document}