### 2018-02-22 reports/initial-proposal/initial-proposal.tex

%        File: initial-proposal.tex
%     Created: Tue Feb 20 10:00 PM 2018 E
% Last Change: Tue Feb 20 10:00 PM 2018 E
%
\documentclass{amsart}
\usepackage{../am221}

\usepackage[margin=1in]{geometry}
\linespread{1.1}

\begin{document}

\title{APMTH 221 Final Project Proposal}
\author{Jiafeng (Kevin) Chen \and Francisco Rivera}
\date{\today}

\maketitle

\section{Collaboration}

Francisco and Kevin will work together to research the relevant optimization
techniques for efficiently computing Wasserstein distances and solving optimal transport problems. Then, they
will split up techniques, and through implementations of them, empirically (and
when possible, theoretically) assess their performance. Time permitting, they
will then attempt to create a new algorithm that performs competitively with
state of the art benchmarks.

\section{Problem Statement}

Given two probability distributions $p_X$ and $p_Y$, the \emph{$p$\textsuperscript{th} Wasserstein distance} is the following optimization problem: \[
W_p(p_X,p_Y) = \pr{\inf_{(X,Y)} \br{ \mathbb{E}_{(X,Y)} \bk{d(X,Y)^p}: X\sim p_X, Y\sim p_Y }}^{1/p},
\]
where the minimum is taken over all joint distributions with marginals $p_X,p_Y$. The Wasserstein distance is a probabilistic view of a class of problems called \emph{optimal transport problems} \cite{peyre2017computational}. If we think of a
probability distribution as a collection of ``mass'' at different points in $M$
(with density of the mass proportional to the probability measure), then the
Wasserstein distance captures the cost of transforming one distribution into the
other in units of mass times distance.  This analogy is particularly
interpretable if we consider discrete probability measures, which are collections of point masses.

Solving the optimization problem underpinning this metric is in general hard and
remains an unsolved problem. We will aim to test the effectiveness of different
optimization techniques to this problem, with particular focus on discrete
distributions.

\section{Motivation}

Numerous important applications exist for the Wasserstein metric. For instance, if we consider images as a discrete probabilistic distribution over $\mathbb{R}^2$,\footnote{For example, we may consider black-and-white images, where the probability mass is proportional to the pixel intensity.} the Wasserstein distance turns out to mimick human perceptions of similarity: Images that are similar to humans are close in Wasserstein distances. Other distances over distributions, such as the well-known Kullback-Leibler divergence, do not have such properties. Thus, Wasserstein distances are extremely important in computer vision, machine learning, and statistics. 
% TODO: what are these important applications?

\section{Data}

As this is a largely theoretical and methodological project, we shall rely on numerical simulation and analysis. We will generate test distributions using simulation
and do not need to concern ourselves with finding real data. When necessary, we
will use standard distributions found in the literature for consistency.

\section{Deliverables}

We will implement optimization techniques to calculate Wasserstein distances and
attempt a novel algorithm that performs competitively. The deliverable will be a survey of recent literature and a report of our contributions.

\section{Next Steps}

First, we will review the current literature and familiarize ourselves with cutting-edge techniques in solving optimal transport problems.

We have the following tentative ideas which we may or may not individually
implement, but from which we will source our  work,

\begin{itemize}

\item We will use linear programs to calculate Wasserstein distances for
discrete metric spaces.\footnote{This is the Kantorovich problem, see \cite{peyre2017computational}} We will explore average complexity for randomly
generated test cases.

\item Training a neural network to hot-start an optimization algorithm for the
discrete case, following recent literature on the creative use of deep learning techniques for difficult optimization problems as in \cite{dutting2017optimal}.

\item Exploring other relevant optimization techniques such as simulated
annealing and genetic algorithms.

\end{itemize}

% TODO: what are we doing next?
\bibliographystyle{plain}

\bibliography{proposal.bib}

\end{document}