### 2018-04-16 reports/notes-ch4/notes-ch4.tex
notes for ch4

\documentclass[11pt,reqno]{amsart}

% Page set-up
\linespread{1.2}

% Packages
\usepackage{../am221}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[capitalise,nameinlink]{cleveref}
\hypersetup{colorlinks=true, citecolor=Blue, linkcolor=Brown}

% Commands
\renewcommand{\b}{\mathbf}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\KL}{\operatorname{KL}}
\newcommand{\Lag}{\operatorname{Lagrange}}

% Thm definitions
\theoremstyle{definition}
\newtheorem{prop}{Proposition}
\theoremstyle{remark}
\newtheorem{rmk}{Remark}

\title{Notes on Entropic Regularization}
\author{Jiafeng Chen\and Francisco Rivera}
\date{\today}

\begin{document}
\maketitle

\section{Statement}

The original problem is \begin{align*}
\min \,& \sum_{i,j} P_{ij} C_{ij}\\
\text{s.t. }&\b P\one = \b a \\
&\b P^T \one = \b b
\end{align*}
\newcommand{\feasible}{\b U(\b a, \b b)}
Denote the feasible region by $\feasible$. \emph{Entropic regularization}
 considers the following modified problem: \[
L^\epsilon_{\b C}(\b a ,\b b) = \min_{\b P \in \feasible} \sum_{i,j} P_{ij}C_
{ij} + \epsilon \underbrace{\sum_
{i,j} P_{ij}
(\log P_{ij} - 1)}_{-H(\b P)}.
\]
We can show that the negative entropy $-H(\b P)$ is $1$-strongly convex (i.e.
 Hessian minus $I$ is
 positive semi-definite). Thus the objective is $\epsilon$-strongly convex. The
 strong convexity means that the optimum $\b P_\epsilon$ is unique.
 Unsurprisingly, $\b P_\epsilon \to \b P^\star$ as $\epsilon \to 0$ where $\b
 P^\star$ is the maximal-entropy solution to the original problem $L_{\b C}^0$.
 Moreover, note that if $\epsilon \to \infty$, we are essentially maximizing the
 entropy, and unsurprisingly, the maximal entropy solution is $\b a \b b^T$:
 In terms of probability, maximal entropy joint distribution is assuming
 independence.

We can reformulate the problem via \emph{KL projection}. Note that minimizing
 the regularized objective is akin to minimizing \begin{align*}
 &\sum_{i,j}\frac{1}{\epsilon}P_{ij}C_{ij} + P_{ij}(\log P_{ij} - 1) + e^{-C_
{ij}/\epsilon} \\
=& \sum_{i,j} P_{ij} \pr{\log P_{ij} - \log\pr{e^{-C_{ij}/\epsilon}}} - P_{ij} +
+ e^{-C_
{ij}/\epsilon} \\
=& \sum_{i,j}P_{ij} \log\frac{P_{ij}}{K_{ij}} - P_{ij} + K_{ij} =: \KL
\binom{\b P}{\b K},
 \end{align*}
where $\b K_{ij} = \exp(-C_{ij}/\epsilon)$ is called a \emph{Gibbs kernel}. Thus
 the problem is akin to projecting $\b K$ onto $\feasible$ via the KL-divergence
 as a distance metric.

The Lagrangian of the regularized problem is \begin{align*}
\Lag(\b P, \b f, \b g) = & \sum_{i,j} P_{ij}C_{ij} + \epsilon \sum_
{i,j} P_{ij}
(\log
 P_{ij} - 1) \\ 
& + \sum_i f_i\pr{a_i - \sum_{j}P_{ij}} + \sum_j g_j \pr{b_j - \sum_i P_{ij}},
\end{align*}
for which the first-order conditions yield \[
C_{ij} + \epsilon \log P_{ij} - f_i - g_j = 0.
\]
Rewriting yields \[
P_{ij} = e^{f_i/\epsilon} e^{-C_{ij} / \epsilon} e^{g_j / \epsilon} = u_i K_{ij}
 v_j
\]
for some $\b u, \b v$. The condition prescribed by $\feasible$, written via $\b
 u, \b v$, yield \[
\b u\odot (\b K \b v) = \b a \quad \b v  \odot(\b K^T \b u) = \b b.
\]
Since the optimum is unique, we need only find $\b u, \b v$ such that these
 conditions hold.

Given $\b v$, we can compute $\b u$ via $\b u = \frac{\b a}{\b K \b v}$ and
 vice versa. This motivates Sinkhorn's algorithm by initializing $\b v^{(0)} =
 \one$ and computing \[
\b u^{(\ell + 1)} = \frac{\b a}{\b K \b v^{(\ell)}} \quad \b v^{(\ell + 1)} = \frac{\b b}
{\b K^T u^{(\ell + 1)}}
\]

It has been shown (2017!) that Sinkhorn updates achieves a $\tau$-approximate
 solution of the unregularized problem in $O(n^2 \log(n)\tau^{-3})$ update
 iterations.

% \bibliographystyle{alpha}
% \bibliography{../optimal-transport}
\end{document}