### 2018-04-25 reports/paper/paper.tex
paper + code

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
\usepackage[preprint]{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{../am221}

% \title{Formatting instructions for NIPS 2018}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

% \author{
%   David S.~Hippocampus\thanks{Use footnote for providing further
%     information about author (webpage, alternative
%     address)---\emph{not} for acknowledging funding agencies.} \\
%   Department of Computer Science\\
%   Cranberry-Lemon University\\
%   Pittsburgh, PA 15213 \\
%   \texttt{hippo@cs.cranberry-lemon.edu} \\
%   %% examples of more authors
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \AND
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
% }

% Commands
\renewcommand{\b}{\mathbf}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\E}{\mathbb{E}}

% Thm definitions
\theoremstyle{definition}
\newtheorem{prop}{Proposition}
\theoremstyle{remark}
\newtheorem{rmk}{Remark}

\title{420}
\author{Jiafeng Chen\\School of Engineering and Applied Sciences\\Harvard University\\\url{jiafengchen@college.harvard.edu}\And Francisco Rivera\\School of Engineering and Applied Sciences\\Harvard University\\\url{frivera@college.harvard.edu}}
\date{\today}

\begin{document}
\maketitle

\section{Introduction and Background}

\subsection{The Optimal Transport Problem}

\subsection{Applications} % (fold)
\label{sub:applications}

% subsection applications (end)

\section{Exact Methods}

\subsection{Linear Programming} % (fold)
\label{sub:linear_programming}

% subsection linear_programming (end)

\subsection{Hungarian Algorithm} % (fold)
\label{sub:hungarian_algorithm}

% subsection hungarian_algorithm (end)

\section{Inexact Methods} % (fold)
\label{sec:inexact_methods}

\subsection{Entropic Regularization and Sinkhorn's Algorithm} % (fold)
\label{sub:entropic_regularization_and_sinkhorn_s_algorithm}

\subsubsection{Convergence}

\subsubsection{Numerical Experiments}

% subsection entropic_regularization_and_sinkhorn_s_algorithm (end)

\subsection{Genetic Algorithms (Hokey)} % (fold)
\label{sub:genetic_algorithms_}

% subsection genetic_algorithms_ (end)
% section inexact_methods (end)

\section{Applications and Experiments} % (fold)
\label{sec:applications_}

% section applications_ (end)

\section{Conclusion} % (fold)
\label{sec:conclusion}

% section conclusion (end)

\bibliographystyle{alpha}
\bibliography{../optimal-transport}
\end{document}