### 2018-04-25 reports/paper/paper.tex
50% of paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
\usepackage[preprint]{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\hypersetup{colorlinks=true, citecolor=Blue, linkcolor=Brown}

\usepackage[capitalise,nameinlink]{cleveref}

\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{../am221}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{bbm}
\usepackage{amsthm}

\newtheorem{theorem}{Theorem}

\linespread{1.1}
% \title{Formatting instructions for NIPS 2018}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

% \author{
%   David S.~Hippocampus\thanks{Use footnote for providing further
%     information about author (webpage, alternative
%     address)---\emph{not} for acknowledging funding agencies.} \\
%   Department of Computer Science\\
%   Cranberry-Lemon University\\
%   Pittsburgh, PA 15213 \\
%   \texttt{hippo@cs.cranberry-lemon.edu} \\
%   %% examples of more authors
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \AND
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
%   %% \And
%   %% Coauthor \\
%   %% Affiliation \\
%   %% Address \\
%   %% \texttt{email} \\
% }

% Commands
\renewcommand{\b}{\mathbf}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\E}{\mathbb{E}}

\newcommand{\KL}{\operatorname{KL}}
% Thm definitions
\theoremstyle{definition}
\newtheorem{prop}{Proposition}
\theoremstyle{remark}
\newtheorem{rmk}{Remark}

\title{Let's Get Movin': Frontiers of Optimal Transport}
\author{Jiafeng Chen\\School of Engineering and Applied Sciences\\Harvard University\\\url{jiafengchen@college.harvard.edu}\And Francisco Rivera\\School of Engineering and Applied Sciences\\Harvard University\\\url{frivera@college.harvard.edu}}
\date{\today}

\newcommand{\feasible}{\b U(\b a, \b b)}

\begin{document}
\maketitle
\section{Introduction and Background}

\subsection{The Optimal Transport Problem and Basic Properties}

Let $\b a \in \R^n_{\ge 0}$ and $\b b \in \R^m_{\ge 0}$ be two vectors of
non-negative real numbers such that $\b a^T \one = \b b^T \one = 1$, where $a_i$
and $b_j$ represents the amount of ``mass'' at different positions. We are
interested in a matrix $\b P \in \R^{n\times m}_{\ge 0}$ where $P_{ij}$
prescribes the amount of mass flowing from $a_i$ to $b_j$. Naturally, we must
have $\sum_j P_{ij} = a_i$ and $\sum_i P_{ij} = b_j$, so that mass is preserved.
Let $\b C \in \R^{n\times m}$ be a \emph{cost matrix}, where $C_{ij}$ represents
the cost of transporting one unit of $a_i$ to $b_j$. The \emph{Kantorovich
optimal transport problem} is the following linear program:
\begin{equation}
\label{eq:ot_problem}
\min_{\b P \in \feasible} \, \sum_{i,j} P_{ij} C_{ij} \tag{Kantorovich}
\end{equation}
where \[
\feasible = \{\b P \in \R^{n\times m}_{\ge 0}: \b P \one = \b a, \b P^T \one =
 \b b\}.
\]

\begin{rmk}[A probabilistic view]
Let $X,Y$ be discrete random variables with support $\b x \in \mathcal X^m$ and
$\b y \in \mathcal Y^n$, such that their marginal distributions are prescribed
by $P(X = x_i) = a_i$ and $P(Y = y_j) = b_j$. Let $P_{ij} = P(X = x_i, Y =
y_j)$, making $\b P = (P_{ij})_ {ij}$ a joint distribution. Let $C: \mathcal X
\times \mathcal Y \to \R$ be a cost function. Then the \eqref{eq:ot_problem} is
equivalent to the following:

\[ \min_{\b P}\, \E_{(X,Y) \sim \b P}\bk{C(X,Y)}, \]

where the expectation is taken over valid joint distributions of $(X,Y)$
respecting the marginal distributions. Note that when $\mathcal X = \mathcal Y =
\R^d$ and $C(x,y) = \norm{x-y}^p$, then the optimal value of the above problem
for $X,Y$ is the $p$-Wasserstein distance between the marginal distributions
$\mu_X,\mu_Y$ (up to a power of $1/p$), for discrete, $d$-dimensional
distributions with finite support.
\end{rmk}

\paragraph{Duality} It can be shown that the dual problem of the above LP takes
 the following form:

\begin{align*}
\max_{(\b f, \b g) \in \R^{m} \times \R^n} \, & \sum_i a_i f_i + \sum_j b_j g_j
\\ 
\text{subjected to } & f_i + g_j \le C_{ij}\, \text{ for all $i,j$}.
\end{align*}

The dual problem has a neat economic intuition. Suppose Kevin needs to transport
$\b a$ to $\b b$ but does not understand how. Francisco, his profit-maximizing
colleague, offers him a deal, where Kevin will pay $f_i$ for Francisco to pick
up a unit of mass at $a_i$ and pay $g_j$ for Francisco to dropoff a unit of mass
at $b_j.$ At minimum, Kevin knows that if any $f_i + g_j > C_ {ij}$, then
Francisco is ripping him off, as transporting one unit from $a_i$ to $b_j$ costs
exactly $C_{ij}$. If, on the other hand, $f_i + g_j \le C_{ij}$ is satisfied,
then given any $\b P$, we have

\[
\sum_{i,j} C_{ij}P_{ij} \ge \sum_{i,j} f_i P_{ij} + \sum_{i,j} g_j P_{ij} =
\sum_i f_i a_i + \sum_j g_j b_j \tag{Weak duality},
\]

which means that Kevin cannot lose by taking Francisco's deal. Strong duality,
which indeed holds, here imply that Francisco's optimal profit is zero, and the
trade is fair.

% Note that, fixing $\b f$, the optimal choice of $g_j$ is
% \[ g_j = \min_i (C_{ij} - f_i). \]
% Let $\b g = \b f^{\b C}$ where $(\b f^{\b C})_j = \min_i (C_{ij} - f_i)$,
% which is called a $\b C$-transform of $\b f$; we may reformulate the dual
% problem in terms of only one choice variable $\b f$.

Given optimal solutions to the primal and dual, $\b P$, $(\b f, \b g)$,
respectively, \emph{complementary slackness} implies that

\[ P_{ij} (C_{ij} - f_i - g_j) = 0 \tag{Complementary slackness} \]

holds, which means that $\b P_{ij} \neq 0 \implies C_{ij} = f_i - g_j$ and
vice versa. 
\subsection{Applications} % (fold)
\label{sub:applications}

% subsection applications (end)

\section{Exact Methods}

\subsection{Linear Programming} % (fold)
\label{sub:linear_programming}

% subsection linear_programming (end)

\subsection{Hungarian Algorithm} % (fold)
\label{sub:hungarian_algorithm}

% subsection hungarian_algorithm (end)

\section{Inexact Methods} % (fold)
\label{sec:inexact_methods}

\subsection{Entropic Regularization and Sinkhorn's Algorithm} % (fold)
\label{sub:entropic_regularization_and_sinkhorn_s_algorithm}

Let \[H(\b P)= -\sum_{i,j} P_{ij}(\log P_{ij} - 1)\] be the \emph{entropy} of
 the transport matrix $\b P$. Observe that \[
\nabla^2_{\b P} (-H(\b P)) - I \text{ is positive semidefinite},
\]
and so $-H(\b P)$ is 1-strongly convex. We regularize the objective of 
\eqref{eq:ot_problem} by $-\epsilon H(\b P):$ \begin{equation}
\label{eq:entropic}
  \min_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
\tag{Entropic Regularization}
\end{equation}
so as to make the objective $\epsilon$-strongly convex. \emph{Entropic
 regularization} approximates \eqref{eq:ot_problem} by solving 
\eqref{eq:entropic}, where \[
\argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
 {\underset{n\to\infty}{\longrightarrow}}
 \argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij},
\]
where the convergence is towards the maximal-entropy solution of 
\eqref{eq:ot_problem}.

Leveraging duality theory, the Lagrangian of the regularized problem is 
\begin{align*}
\mathcal L(\b P, \b f, \b g) = & \sum_{i,j} P_{ij}C_{ij} + \epsilon \sum_
{i,j} P_{ij}
(\log
 P_{ij} - 1) \\ 
& + \sum_i f_i\pr{a_i - \sum_{j}P_{ij}} + \sum_j g_j \pr{b_j - \sum_i P_{ij}},
\end{align*}
where the first-order condition on $\b P$ shows that the optimum $\b P^\*$
 satisfies 
\begin{equation}
\label{eq:foc}
  C_{ij} + \epsilon \log P_{ij}^\* - f_i^\star - g_j^\* = 0 \implies \b P^\* =
 \diag(\exp(\b f^\* / \epsilon)) \exp(-\b C/\epsilon )\diag(\exp(\b g^\* /
 \epsilon))
\end{equation}
for some $\b f^\star, \b g^\*,$ where the $\exp$ is taken component-wise.
 Suppose $\b f^\star, \b g^\*$ is such that $\b P^\star$ defined by 
\eqref{eq:foc} is within the constraints $\feasible$, then the value of the
 Lagrangian $\mathcal L(\b P^\*, \b f^\*, \b g^\*)$ equals the value of 
\eqref{eq:entropic}, and since  $\b P^\*$ is chosen to maximize the strongly
 concave function $\b P \mapsto \mathcal L(\b P, \b f^\*, \b g^\*)$, we have
 that $(\b P^\*, \b f^\*, \b g^\*)$ solves both programs.

Before proceeding, we define some notation that is a hybrid of 
\cite{peyre2017computational} and \cite{altschuler2017near}. Let $\b A = \exp
(-\b C / \epsilon),$ $\b u = \exp(\b f / \epsilon)$, $\b v = \exp(\b g /
 \epsilon)$, $\b x = \log \b u$, $\b y = \log \b v$, $\b X = \diag(\b u)$, $\b
 Y = \diag(\b v)$, where $\exp, \log$ are taken componentwise. Thus, we aim to
 find vectors $\b u, \b v$, such that \[
\b P^\* = \b X \b A \b Y.
\]
satisfies the constraints $\feasible.$ In terms of $\b u, \b v$, the constraints
 are \begin{equation}
\label{eq:sinkhorn}
   \b u \odot (\b A \b v) = \b a \quad \text{    and    } \quad \b v \odot (\b A^T
 \b u) = \b b 
 \end{equation}
\eqref{eq:sinkhorn} shows that, given $\b v$, we can only set $\b u = \frac{\b
 a }{\b A \b v}$, where division is componentwise. This motivates the
 Sinkhorn-Knopp algorithm \cite{sinkhorn1967concerning}, which iteratively
 updates
 $\b u, \b v$ such that one of the constraints hold exactly in this fashion. In
 practical implementation, to prevent numerical underflow, the algorithm should
 work in log-space with $\b x, \b y$. \cref{alg:sinkhorn} presents a numerically
 stable implementation of Sinkhorn-Knopp that terminates when the $\ell_1$ error
 on constraint violation gets sufficiently small. At each step of 
\cref{alg:sinkhorn}, if $k$ is odd, then $\b P^{(k)}$ matches row-wise with $\b
 a$, and if $k$ is even, then $\b P^{(k)}$ matches column-wise with $\b
 b$.

\begin{algorithm}
\caption{Numerically stable Sinkhorn-Knopp algorithm as presented in 
\cite{altschuler2017near}}
\label{alg:sinkhorn}

\begin{algorithmic}

\Function{Sinkhorn}{$\b A, \b a, \b b, \epsilon$}
\State Initialize $k=0$
\State $\b P^{(0)} \gets \b A, \b x^0 \gets 0, \b y^0 \gets 0$
\While{$\norm{\b P^{(k)}\one - \b a}_1 + \norm{(\b P^{(k)})^T\one - \b b}_1 >
 \epsilon$}
\State $k\gets k + 1$
\If{$k$ is odd}
\State $\Delta \b x \gets \log\frac{\b a}{\b P^{(k-1)}\one}$
\State $\b x^k \gets \b x^{k-1} + \Delta \b x, \b y^k \gets \b y^{k-1}$
\EndIf
\If{$k$ is even}
\State $\Delta \b y \gets \log\frac{\b b}{(\b P^{(k-1)})^T\one}$
\State $\b x^k \gets \b x^{k-1}, \b y^k \gets \b y^{k-1}  + \Delta \b y$
\EndIf
\State $\b P^{(k)} \gets \b X^k \b A \b Y^{k}$
\EndWhile
\State\Return{$\b P^{(k)} / \norm{\b P^{(k)}}_1$}
\Comment{$\b P^{(0)}$ is not normalized; $\b P^{(k)}, k > 0$ is normalized.}
\EndFunction
\end{algorithmic}
\end{algorithm}

Note that \cref{alg:sinkhorn} does not output a matrix that satisfies the
 conditions $\feasible$ exactly. However, \cite[Algorithm 2]{altschuler2017near}
 analyzes a rounding scheme that forces the output to be inside $\feasible$. The
 rounding scheme introduces error on the order of $\epsilon$ and thus does not
 pose any problems for solving the original program \eqref{eq:entropic}. The
 rest of this section will be devoted to an analysis of \cref{alg:sinkhorn}.

\subsubsection{Convergence}
\emph{A priori}, it is unclear that \cref{alg:sinkhorn} converges, nor is the
 speed of convergence obvious. 
\cite{altschuler2017near} presents an extremely succinct analysis of 
\cref{alg:sinkhorn}, which we reproduce in this section. Let \[
f(\b x, \b y) = \sum_{i,j} A_{ij} e^{x_i + y_j} - \sum_i a_i x_i - \sum_j b_j
 y_j.
\] 
The first order conditions for the optimality of $f$ gives $\b x, \b y$
 that ensures \eqref{eq:sinkhorn}. Sinkhorn-Knopp can be viewed as coordinate
 descent on $f$: To optimize $x_i^k$, the first order condition yields \[
x_i^k = \log \frac{a_i}{\sum_j A_{ij} e^{y_j^{k-1}}} = \log \frac{a_i}{e^
{x_i^{k-1}}\sum_j A_{ij} e^{y_j^{k-1}}} + x_i^{k-1} = \log \frac{a_i}{\b P^{
(k-1)}\one} + x_i^{k-1},
\]
which is exactly the update rule in \cref{alg:sinkhorn}.

The proof of \cite{altschuler2017near} uses the follow strategy. First, we
 calculate the step gain in the Sinkhorn-Knopp algorithm for optimizing $f$ in
 terms of Kullback-Leibler divergences.
 Next we bound the distance of $f$'s initial value from its optimal value.
 Lastly, using Pinsker's inequality, we show that if \cref{alg:sinkhorn} does
 not terminate, then the step size must be larger than $C\epsilon^{2}$, which
 then gives a bound of convergence in $\epsilon^{-2}$.

For $k > 1$, we can
 calculate the gain of each step \cite[Lemma 2]{altschuler2017near}:
\begin{align}
f(\b x^{k-1}, \b y^{k-1}) - f(\b x^{k}, \b y^{k}) &= \pr{\sum_{i,j} A_{ij}e^
{x_i^{k-1} + y_i^{k-1}} - \sum_{i,j} A_{ij}e^
{x_i^{k} + y_i^{k}}} \nonumber\\&\qquad+ \ip{\b a, \b x^k -\b x^{k-1}} + \ip{\b
 b,
 \b y^k - \b y^{k-1}}
 \nonumber\\
&= (1-1) + \KL(\b a || \b P^{(k)}\one) + \KL(\b b || (\b P^{(k)})^T\one)
 \nonumber \tag{Writing out the inner products}\\
&= \KL(\b a || \b P^{(k)}\one) + \KL(\b b || (\b P^{(k)})^T\one). 
\label{eq:klsum}
\end{align}
\eqref{eq:klsum} gives an elegant formulation of the gain each step exactly as a
 function of the Kullback-Leibler divergence between the target distribution and
 the current marginal distribution.

Next, we claim that the total distance in $f$ that we are required to
 travel is bounded above \cite[Lemma 3]{altschuler2017near}. Note that 
\cref{alg:sinkhorn} behaves almost identically---producing the same $\b P^{(k)}$
 at each iteration---if the input $\b A$ passed in is
 multiplied by a constant. Thus we may without loss of generality replace $\b A$
 with
 $\b A_0 = \b A / \norm{\b A}_1$. Let $f^k$ denote $f(\b x^k, \b y^k)$ and
 $f^\*$ denote $\min f$. Then we have $
f(0,0) - f^1 = \KL(\b a || \b A_0\one) \ge 0.
$
Thus \[
f^1 - f^\* \le f(0,0) - f^\* = \sum_{i} a_i x_i^\* + \sum_j b_j y_j^\* \le \max_
{i,j}
(x_i^\* + y_j^\*).
\]
Observe that $
A_{0, ij}e^{x_i^\* + y_j^\*} \le \sum  A_{0, ij}e^{x_i^\* + y_j^\*} = 1.
$
Therefore $x_i^\* + y_j^\* \le -\log A_{0,ij} \le \log \frac{\norm{\b A}_1}{\min_
{i,j} A_{ij}}$. Putting these together shows that \begin{equation}
\label{eq:total}
  f^1 - f^\* \le \log \frac{\norm{\b A}_1}{\min_
{i,j} A_{ij}}.
\end{equation}

Lastly, Pinsker's inequality \cite[Lemma 4]{altschuler2017near}, a well-known
 result in information theory, states that for any probability measures $p, q$,
 \[
\norm{p-q}_1^2 \le 2 \KL(p||q). 
\]
Applying Pinsker's inequality shows that every step in \cref{alg:sinkhorn}
 before termination has \[
\epsilon < \norm{\b P^{(k)}\one - \b a}_1 + \norm{(\b P^{(k)})^T\one - \b b}_1
 \le \bk{4\pr{\KL(\b a || \b P^{(k)}\one) + \KL(\b b || (\b P^{(k)})^T\one)}}^
{1/2}.
\]
Therefore we improve $f$ by more than $\frac{1}{4}\epsilon^2$ every step before
 termination. Since the total improvement is bounded by \eqref{eq:total}, we
 arrive at the main result of \cite{altschuler2017near}:
\begin{theorem}[\cite{altschuler2017near}]
\label{thm:sinkhorn_convergence}
  \cref{alg:sinkhorn} terminates in at most $4\epsilon^{-2}\log \frac{\norm{\b A}_1}{\min_
{i,j} A_{ij}}$ steps.
\end{theorem}

\begin{figure}[tb]
  \centering
  \includegraphics{figs/sinkhorn.eps}
  \caption{Numerical experiments on the speed of $\ell_1$ convergence for \cref{alg:sinkhorn}}
  \label{fig:sinkhorn_convergence}
\end{figure}

\subsubsection{Numerical Experiments}
\label{sub:numerical_sink}
We implement \cref{alg:sinkhorn} and compute the $\ell_1$ distance $\norm{\b P^{
(k)}\one - \b a}_1 + \norm{(\b P^{(k)})^T\one - \b b}_1$ as a function of $k$.
 We generate random marginal distribution targets $\b a, \b b$ by sampling
 from a uniform distribution on the unit interval and normalizing by the sum. We
 generate random cost matrices by sampling each component from a uniform
 distribution. We plot the log-log plot in \cref{fig:sinkhorn_convergence}. Note
 that if the results of \cref{thm:sinkhorn_convergence} is tight for these
 random matrices and constraints, then the log-log plot should look linear with
 slope $-2$. This is not what we observe in \cref{fig:sinkhorn_convergence},
 suggesting that the average case---if not worst case---of \cref{alg:sinkhorn}
 may be better than $O(\epsilon^{-2})$. In fact, if we plot the log of error
 against $k$, we find that the relationship is linear, suggesting that, for these
 random matrices and distributions, the behavior of \cref{alg:sinkhorn} is
 $O(\log(1/\epsilon))$.

% subsection entropic_regularization_and_sinkhorn_s_algorithm (end)

\subsubsection{Towards a Tightness Analysis of \cref{thm:sinkhorn_convergence}}
\cref{sub:numerical_sink} suggests that the bound of $\epsilon^{-2}$ may be far
 too generous. In this section we sketch a potential direction in analyzing the
 worst-case performance of \cref{alg:sinkhorn}. \cite{sason2015reverse} gives a
 reverse Pinsker's inequality: For discrete probability measures on finite
 support $p,q$, \[
\KL(p||q) \le \log \pr{1+\frac{1}{q_{\min}}\norm{p-q}_1^2} \le \frac{1}{q_{\min}}\norm{p-q}_1^2,
\]
where $q_{\min}$ is the minimum of $q$ on its support. The inequality suggests
 that before termination and after the $\ell_1$ error reaches some $\delta$,
 every
 step of \cref{alg:sinkhorn} improves $f$ by at
 most \[
\KL(\b a || \b P^{(k)}\one) + \KL(\b b || (\b P^{(k)})^T\one) \le \frac{C}
{\min(\b P^{(k)}\one, (\b P^{(k)})^T\one)}\delta^2,
\]
for some constant $C$, where the minimum is taken element-wise. We can remove
 the dependence on $k$ of
 the right-hand side by the following consideration. Let $k'$ be the first time
 that the $\ell_1$-error is below $\frac{1}{2}\min(\b a, \b b)$. Then for every
 $k > k'$, $\min(\b P^{(k)}\one, (\b P^{(k)})^T\one) \ge \frac{1}{2}\min(\b a,
 \b b)$, and so the bound becomes \begin{equation}
   \label{eq:bound_improvement}
\KL(\b a || \b P^{(k)}\one) + \KL(\b b || (\b P^{(k)})^T\one) \le \frac{C'}
{\min(\b a,
 \b b)}\delta^2,
 \end{equation}
for $k > k'$. Let $k_\epsilon$ be the step immediately before termination. Let
 $k_\delta$ be the first iteration such that $k_\delta > k'$ and that $k_\delta$
 reaches $\ell_1$-error of $\delta$.
If we
 can show that there is some inputs $\b A, \b a, \b
 b,\epsilon$ for
 which we can bound the total improvement $f^{k_\eta} - f^{k_\epsilon}$ below by
 some quantity $M(\b A, \b a, \b
 b,\epsilon, \delta)$, then for the inputs $\b A, \b a, \b
 b,\epsilon$, \cref{alg:sinkhorn} has a lower bound of $\Omega\pr{M(\b A, \b a,
 \b
 b,\epsilon, \delta)\delta^{-2}\min(\b a, \b b)}$. Here, $\delta$ should be
 implicitly a function of $\epsilon$, e.g. $\delta = \sqrt{\epsilon}$, so that
 the bound is a function of $\epsilon$. Choosing $\delta(\epsilon)$ such that
 the bound $M(\b A, \b a,
 \b
 b,\epsilon, \delta)$ is tractable is a major difficulty, which we leave to
 future work.

\subsection{Genetic Algorithms (Hokey)} % (fold)
\label{sub:genetic_algorithms_}

% subsection genetic_algorithms_ (end)
% section inexact_methods (end)

\section{Applications and Experiments} % (fold)
\label{sec:applications_}

% section applications_ (end)

\section{Conclusion} % (fold)
\label{sec:conclusion}

% section conclusion (end)

\bibliographystyle{alpha}
\bibliography{../optimal-transport}
\end{document}